Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-03-06 paper/current/main.tex
Plot prestige rank with and without included axis points

The effective distance  estimated by our embedding model explains and predicts scientific mobility best. 
Embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds for the variance of domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d). 
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance globally (Fig.~\ref{fig:gravity_performance}c--d) and for domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility(Fig.~\ref{fig:supp:gravity_correlation}g--h).
Similarly, embedding ``diffusion'' distance outperforms distances derived from calculating personalized-page rank scores on the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}). 
Geographic distance alone is a poor tool for understanding international scientific mobility---rather, by learning a distance directly from the data, it is possible to better explain, and better predict flux between organizations.

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms and is more robust than geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) explains more of the expected flux between organization pairs than does geographic distance (\textbf{b}). 
		Red line is line of best fit. 
		Black dots are mean flux across binned distances.
		Color indicates frequency of organization pairs within each hex bin.
		\textbf{c-d.} Predictions  of flux between organization pairs made using embedding distance outperform those made using geographic distance. 
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $x = y$.
		\textbf{e-f.} Predictions made using model parameters learned from subset of same-country organization pairs. 
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only pairs of organizations in different countries. 
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets .
	}
	\label{fig:gravity_performance}
\end{figure}

@@

%
% Figure - SemAxis results
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/SemAxis/semaxis_compare_fig.pdf}
	\caption{
		\textbf{SemRank hierarchy is robust.}
		\textbf{a.} Spearman correlation (n = 143) between Times prestige rank and SemAxis rank, by the number of high and bottom-ranked (geographically matched) universities aggregated to form the poles. 
		More universities improves performance, but quickly saturates after around five universities. 
		\textbf{b - f.} Comparison between the Times and SemAxis ranks of universities, by number of universities used to define the poles (n). 
		White points are those top and bottom 20 universities aggregated to define the ends of the axis. 
		The grey box corresponds to the top 20 and bottom 20 ranks. 
	  	PCC details the estimate from Spearman correlation between the two rankings. 
	}
	\label{fig:supp:semaxis_compare}
\end{figure}

>>>>

The effective distance estimated by our embedding model explains and predicts scientific mobility best.
The amount of variance explained by embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds for the variance of only domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d).
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance globally (Fig.~\ref{fig:gravity_performance}c--d) and for domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility(Fig.~\ref{fig:supp:gravity_correlation}g--h).
Similarly, embedding distance outperforms ``diffusion''  distances such as those derived from calculating personalized-page rank scores on the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
Geographic distance alone is a poor tool for understanding international scientific mobility---rather, by learning a distance directly from the data, it is possible to better explain, and better predict flux between organizations.

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms and is more robust than geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) explains more of the expected flux between organization pairs than does geographic distance (\textbf{b}).
		Red line is line of best fit.
		Black dots are mean flux across binned distances.
		Error bars are 99\% confidence intervals.
		Color indicates frequency of organization pairs within each hex bin.
		\textbf{c-d.} Predictions  of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $x = y$.
		``Err'' is calculated as the root-mean squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets .
	}
	\label{fig:gravity_performance}
\end{figure}

@@

%
% Figure - SemAxis results
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/SemAxis/semaxis_compare_fig.pdf}
	\caption{
		\textbf{SemRank hierarchy is robust.}
		\textbf{a.} Spearman correlation (n = 143) between Times prestige rank and SemAxis rank, by the number of high and bottom-ranked (geographically matched) universities aggregated to form the poles.
		Black points show spearman correlation using all organizations; white points show correlation using only universalizes not aggregated in the poles. 
		More universities improves performance, but quickly saturates after around five universities.
		\textbf{b - f.} Comparison between the Times and SemAxis ranks of universities, by number of universities used to define the poles (n).
		White points are those top and bottom 20 universities aggregated to define the ends of the axis.
		The grey box corresponds to the top 20 and bottom 20 ranks.
	  	PCC details the estimate from Spearman correlation between the two rankings.
	}
	\label{fig:supp:semaxis_compare}
\end{figure}
